{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or, How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-03-04 thesis/landslide.tex
few revisions dpor

{\bf Identifying new interleavings.}
To summarize, for each pair of transitions $t_i$ and $t_j$ ($i<j$),
DPOR identifies interleavings which will reorder them,
executing $t_j' t_i'$ to reach a new program state not exposed in the current interleaving.
If $t_i \rightarrow_S t_j$ then they cannot be reordered,
and if $t_i \not\leftrightsquigarrow_S t_j$ then reordering them
will produce a state already encountered in this interleaving;
hence, DPOR considers only transition pairs
which are not related by happens-before and are related by memory conflict ({\tt is\_evil\_ancestor()}).

% TODO: change i<4 to i<j.
% i accidentally typed i<4 here and decided it'd be funny to leave it this way
For each such pair $t_i$, $t_j$, $i<4$, let $s$ denote the state (preemption point) immediately before $t_i$.
Then:
\begin{itemize}
	\item If $T(t_j)$ is runnable at $s$, return $I_{ij} = (t_1 \dots t_{i-1}, T_j)$ ({\tt tag\_good\_sibling()}).
	\item Otherwise, there must be some third thread runnable at $s$
		(from $t_i \not\rightarrow_S t_j$, by contrapositive of the second condition of dependence);
		then, return all $I_{ik} = (t_1 \dots t_{i-1}, T_k)$ such that
		$T_k \ne T(t_i)$ and $T_k$ is runnable at $s$
		({\tt tag\_all\_siblings()}).
\end{itemize}

@@

Acknowledging that quickly gaining an intuitive understanding for an algorithm
is basically impossible from just paragraphs of text,
I provide in this section an example of a single iteration of DPOR
on the example state space from Figures~\ref{fig:concurrency-bug} and \ref{fig:tree}.

@@

Whenever a memory conflict is identified for DPOR as described above,
the access pair's corresponding locksets and/or happens-before edges are checked to determine if it's also a data race.
Note the distinction: DPOR memory conflicts indicate that two thread transitions,
if reordered, could produce different behaviour, even if all accesses therein are adequately synchronized;
while a data race indicates furthermore that the two threads can be interleaved precisely at the moment of one or both accesses.
% nb. i think "only one of the accesses can be preempted at and have the other interleaved inside, but not the other way around,
% is a property of cli/sti involvement in kernel space only.

@@

Conditions 1, 2, and 4 defined in \sect{\ref{sec:background-hb}}, provided the Limited Happens-Before definition for \#4,
coincide exactly with the conditions to be considered a memory conflict under DPOR.
Hence all that remains to be checked is \#3, the set of locks held by each thread at the time of access.

>>>>

{\bf Identifying new interleavings.}
To find which alternate interleavings need to be tested,
% TODO: change i<4 to i<j.
% i accidentally typed i<4 here and decided it'd be funny to leave it this way
DPOR compares pairwise each pair of transitions $t_i$ and $t_j$, $i<4$, in the current interleaving $S$.
If $t_i \rightarrow_S t_j$ then they cannot be reordered,
and if $t_i \not\leftrightsquigarrow_S t_j$ then reordering them
will produce a state already encountered in this interleaving;
hence, DPOR marks new interleavings only
when $t_i \not\rightarrow_S t_j$ and $t_i \leftrightsquigarrow_S t_j$
%for transition pairs not related by happens-before and are related by memory conflict
({\tt is\_evil\_ancestor()}).
\footnote{To aid intuition, consider the two extremes:
if all transitions are related by happens-before,
the program is not concurrent and no alternate interleavings are possible;
if all transitions are memory-independent,
the program exhibits full data isolation between threads and all schedules are observably equivalent.}

For each such pair, let $s$ denote the state (preemption point) before $t_i$.
Then:
\begin{itemize}
	\item If $T(t_j)$ is runnable at $s$, return $I_{ij} = (t_1 \dots t_{i-1}, T_j)$ ({\tt tag\_good\_sibling()}).
	\item Otherwise, there must be some third thread runnable at $s$;
		\footnote{AFSOC $T(t_i)$ is the only runnable thread at $s$,
		then either $T(t_i)$'s execution at $s$ enabled $T(t_j)$,
		or it enabled an intermediate transition
		(whether by $T(t_i)$ or a third thread)
		which in turn enabled $T(t_j)$.
		In either case $t_i$ is transitively dependent with $t_j$, contradicting $t_i \not\rightarrow_S t_j$.}
		%(from $t_i \not\rightarrow_S t_j$, by contrapositive of the second condition of dependence);
		then, return all $I_{ik} = (t_1 \dots t_{i-1}, T_k)$ such that
		$T_k \ne T(t_i)$ and $T_k$ is runnable at $s$
		({\tt tag\_all\_siblings()}).
\end{itemize}

To summarize, %for each pair of transitions $t_i$ and $t_j$ ($i<j$) in $S$,
DPOR identifies $I_{ij}$s which will (eventually) reorder each conflicting, concurrent transition pair in $S$
%executing some $t_j' t_i'$
to reach a (possibly) new program state not exposed in the current interleaving.

@@

Acknowledging that quickly gaining an intuitive understanding for an algorithm
is basically impossible from just paragraphs of text,
I conclude this section with an example of a single iteration of DPOR
on the example state space from Figures~\ref{fig:concurrency-bug} and \ref{fig:tree}.

@@

Whenever a memory conflict is identified for DPOR as described above,
the access pair's corresponding locksets and/or happens-before edges are checked to determine if it's also a data race.
Note the distinction: DPOR memory conflicts indicate that two thread transitions,
if reordered, could produce different behaviour, even if all accesses therein are adequately synchronized;
while a data race indicates furthermore that the two threads can be interleaved precisely at the moment of one or both accesses,
supposing that a new preemption point were introduced to split one or both transitions in half.
% nb. i think "only one of the accesses can be preempted at and have the other interleaved inside, but not the other way around,
% is a property of cli/sti involvement in kernel space only.

@@

Conditions \#1, \#2, and \#4 defined in \sect{\ref{sec:background-hb}},
provided the Limited Happens-Before definition for \#4,
coincide with DPOR's version of happens-before described in the previous section.
%coincide exactly with the conditions to be considered a memory conflict under DPOR.
Hence all that remains to be checked is \#3, the set of locks held by each thread at the time of access.
